\textbf{The Mint Scale: A Fresh Validation of the Multimodal Interoception Questionnaire and Comparison to the MAIA, BPQ and IAS}

### 2025-08-20 paper/manuscript.qmd
update

These differing philosophies have practical consequences, leading to poor convergence between scales that are assumed to measure the same construct [@vig2022questionnaires] - a classic example of the "jingle fallacy" where the shared label of "interoception" can mask underlying measurand differences, hindering replication and theoretical integration. In acknowledgement of this issue, efforts have been made to align subjective measures with theoretical models that outline different facets of interoception (e.g., accuracy, attention, deficits). However, focusing on developing measures that are theory-driven may however be premature when theoretical models are in flux, and when a topographic mapping to self-report dimensions is not guaranteed. For example,  interoceptive distinctions that are relevant at a neuroanatomical or cognitive level might not necessarily map onto the structure of subjective experiences. More fundamentally, existing interoception questionnaires typically fail to consider the full range of interoceptive modalities and the context in which interoceptive experiences occur. Most existing questionnaires focus on a narrow range of bodily modalities, with a heavy bias towards cardio-respiratory sensations at the expense of others like gastric, urogenital, or thermoregulatory signals. Perhaps most critically, existing measures ask participants to make general assessments across most/all contexts, which does not capture the potential homeostatic relevance of attending or detecting internal signals. Without a contextual reference point, there is likely to be substantial variability in how participants interpret and respond to items. For example, a general item about "feeling your heart beat" is ambiguous since its salience, frequency, experience and interpretation would differ depending on the context, e.g., on whether it occurs during physical exercise, a moment of anxiety, or quiet rest (e.g., someone who exercises regularly, or is anxious, might more easily recall an instance of occurrence of feeling their heart beating and respond based on that contextual saliency).

To address these gaps, we developed and validated the Multimodal Interoception Questionnaire (Mint), a new questionnaire that systematically addresses the above limitations. In Study 1, we describe the bottom-up development of the Mint, using a "modality-by-context" framework to generate a comprehensive item pool, which was then refined using cutting-edge psychometric network analysis. In Study 2, we validated the final 33-item version of the Mint against the established interoception questionnaires, demonstrating its robust structure and predictive validity for a range of clinically-relevant outcomes.

@@

We would like to thank the dissertation students from the University of Sussex for their help in data collection.
DM would also like to thank Anastasia for the motivation provided to write this paper.

>>>>

These differing philosophies have practical consequences, leading to poor convergence between scales that are assumed to measure the same construct [@vig2022questionnaires] - a classic example of the "jingle fallacy" where the shared label of "interoception" can mask underlying measurand differences, hindering replication and theoretical integration. In acknowledgement of this issue, efforts have been made to align subjective measures with theoretical models that outline different facets of interoception (e.g., accuracy, attention, deficits). However, focusing on developing measures that are theory-driven may however be premature when theoretical models are in flux, and when a topographic mapping to self-report dimensions is not guaranteed. For example,  interoceptive distinctions that are relevant at a neuroanatomical or cognitive level might not necessarily map onto the structure of subjective experiences. More fundamentally, existing interoception questionnaires typically fail to consider the full range of interoceptive modalities and the context in which interoceptive experiences occur. Most existing questionnaires focus on a narrow range of bodily modalities, with a heavy bias towards cardio-respiratory sensations at the expense of others like gastric, urogenital, or thermoregulatory signals.

Perhaps most critically, existing measures often ask participants to make general assessments across most/all contexts, which does not capture the potential homeostatic relevance of attending or detecting internal signals. For example, the BPQ asks participants to reflect on how often during most situations a person is aware of various bodily sensations, such as their heart beating.  Without a contextual reference point, there is likely to be substantial variability in how participants interpret and respond to items as well as what those responses might mean about their interoceptive abilities.  A general item about "feeling your heart beat" is ambiguous since its salience, frequency, experience and interpretation would differ depending on the context, which may or may not be relevant homeostatically. For instance, awareness of one’s heartbeat may occur across multiple contexts, including during physical exercise, a moment of anxiety, excitement, or quiet rest with awareness across different contexts reflecting something different about interoceptive sensations. This is further complicated by retrospective memory biases such that individuals respond with the most easily accessible or recent instances of awareness; for example,  someone who exercises regularly, or is anxious, might more easily recall an instance of occurrence of feeling their heart beating and respond based on that contextual saliency. As an exception, the MAIA does provide some contextual cues for participants (e.g., “When something is wrong in my life I can feel it in my body”) but this is inconsistent across items and focuses almost exclusively on negative (e.g., painful) and emotional (e.g., angry, joyful) contexts as well as general ‘body’ sensations rather than specific interoceptive modalities. Thus, no existing scale captures subjective assessments across multiple interoceptive modalities and across different contexts in which interoceptive experiences occur.

To address this gap, we developed and validated the Multimodal Interoception Questionnaire (Mint), a new questionnaire that systematically addresses the above limitations. In Study 1, we describe the bottom-up development of the Mint, using a "modality-by-context" framework to generate a comprehensive item pool, which was then refined using cutting-edge psychometric network analysis. In Study 2, we validated the final 33-item version of the Mint against the established interoception questionnaires, demonstrating its robust structure and predictive validity for a range of clinically-relevant outcomes.

@@

We would like to thank Hugo Critchley for his valuable feedback on the manuscript.
DM would also like to thank Anastasia for the motivation provided to write this paper.

### 2025-08-20 paper/manuscript.tex
update

\authornote{\par{\addORCIDlink{Dominique
Makowski}{0000-0001-5375-9967}}\par{\addORCIDlink{Ana
Neves}{0009-0006-0020-7599}}\par{\addORCIDlink{Giulia
Poreiro}{0000-0002-2343-5109}} 
\par{ }
\par{     \begin{tcolorbox}[enhanced jigsaw, breakable, opacityback=0, rightrule=.15mm, colback=white, toprule=.15mm, arc=.35mm, bottomrule=.15mm, colframe=quarto-callout-note-color-frame, left=2mm, leftrule=.75mm]

@@

These differing philosophies have practical consequences, leading to
poor convergence between scales that are assumed to measure the same
construct (\citeproc{ref-vig2022questionnaires}{Vig et al., 2022}) - a
classic example of the ``jingle fallacy'' where the shared label of
``interoception'' can mask underlying measurand differences, hindering
replication and theoretical integration. In acknowledgement of this
issue, efforts have been made to align subjective measures with
theoretical models that outline different facets of interoception (e.g.,
accuracy, attention, deficits). However, focusing on developing measures
that are theory-driven may however be premature when theoretical models
are in flux, and when a topographic mapping to self-report dimensions is
not guaranteed. For example, interoceptive distinctions that are
relevant at a neuroanatomical or cognitive level might not necessarily
map onto the structure of subjective experiences. More fundamentally,
existing interoception questionnaires typically fail to consider the
full range of interoceptive modalities and the context in which
interoceptive experiences occur. Most existing questionnaires focus on a
narrow range of bodily modalities, with a heavy bias towards
cardio-respiratory sensations at the expense of others like gastric,
urogenital, or thermoregulatory signals. Perhaps most critically,
existing measures ask participants to make general assessments across
most/all contexts, which does not capture the potential homeostatic
relevance of attending or detecting internal signals. Without a
contextual reference point, there is likely to be substantial
variability in how participants interpret and respond to items. For
example, a general item about ``feeling your heart beat'' is ambiguous
since its salience, frequency, experience and interpretation would
differ depending on the context, e.g., on whether it occurs during
physical exercise, a moment of anxiety, or quiet rest (e.g., someone who
exercises regularly, or is anxious, might more easily recall an instance
of occurrence of feeling their heart beating and respond based on that
contextual saliency).

To address these gaps, we developed and validated the Multimodal
Interoception Questionnaire (Mint), a new questionnaire that
systematically addresses the above limitations. In Study 1, we describe
the bottom-up development of the Mint, using a ``modality-by-context''
framework to generate a comprehensive item pool, which was then refined
using cutting-edge psychometric network analysis. In Study 2, we
validated the final 33-item version of the Mint against the established
interoception questionnaires, demonstrating its robust structure and
predictive validity for a range of clinically-relevant outcomes.

>>>>

\authornote{\par{\addORCIDlink{Dominique
Makowski}{0000-0001-5375-9967}}\par{\addORCIDlink{Ana
Neves}{0009-0006-0020-7599}}\par{\addORCIDlink{Giulia
Poreiro}{0000-0002-2343-5109}} 
\par{ }
\par{     \begin{tcolorbox}[enhanced jigsaw, colframe=quarto-callout-note-color-frame, rightrule=.15mm, colback=white, toprule=.15mm, breakable, left=2mm, opacityback=0, leftrule=.75mm, bottomrule=.15mm, arc=.35mm]

@@

These differing philosophies have practical consequences, leading to
poor convergence between scales that are assumed to measure the same
construct (\citeproc{ref-vig2022questionnaires}{Vig et al., 2022}) - a
classic example of the ``jingle fallacy'' where the shared label of
``interoception'' can mask underlying measurand differences, hindering
replication and theoretical integration. In acknowledgement of this
issue, efforts have been made to align subjective measures with
theoretical models that outline different facets of interoception (e.g.,
accuracy, attention, deficits). However, focusing on developing measures
that are theory-driven may however be premature when theoretical models
are in flux, and when a topographic mapping to self-report dimensions is
not guaranteed. For example, interoceptive distinctions that are
relevant at a neuroanatomical or cognitive level might not necessarily
map onto the structure of subjective experiences. More fundamentally,
existing interoception questionnaires typically fail to consider the
full range of interoceptive modalities and the context in which
interoceptive experiences occur. Most existing questionnaires focus on a
narrow range of bodily modalities, with a heavy bias towards
cardio-respiratory sensations at the expense of others like gastric,
urogenital, or thermoregulatory signals.

Perhaps most critically, existing measures often ask participants to
make general assessments across most/all contexts, which does not
capture the potential homeostatic relevance of attending or detecting
internal signals. For example, the BPQ asks participants to reflect on
how often during most situations a person is aware of various bodily
sensations, such as their heart beating. Without a contextual reference
point, there is likely to be substantial variability in how participants
interpret and respond to items as well as what those responses might
mean about their interoceptive abilities. A general item about ``feeling
your heart beat'' is ambiguous since its salience, frequency, experience
and interpretation would differ depending on the context, which may or
may not be relevant homeostatically. For instance, awareness of one's
heartbeat may occur across multiple contexts, including during physical
exercise, a moment of anxiety, excitement, or quiet rest with awareness
across different contexts reflecting something different about
interoceptive sensations. This is further complicated by retrospective
memory biases such that individuals respond with the most easily
accessible or recent instances of awareness; for example, someone who
exercises regularly, or is anxious, might more easily recall an instance
of occurrence of feeling their heart beating and respond based on that
contextual saliency. As an exception, the MAIA does provide some
contextual cues for participants (e.g., ``When something is wrong in my
life I can feel it in my body'') but this is inconsistent across items
and focuses almost exclusively on negative (e.g., painful) and emotional
(e.g., angry, joyful) contexts as well as general `body' sensations
rather than specific interoceptive modalities. Thus, no existing scale
captures subjective assessments across multiple interoceptive modalities
and across different contexts in which interoceptive experiences occur.

To address this gap, we developed and validated the Multimodal
Interoception Questionnaire (Mint), a new questionnaire that
systematically addresses the above limitations. In Study 1, we describe
the bottom-up development of the Mint, using a ``modality-by-context''
framework to generate a comprehensive item pool, which was then refined
using cutting-edge psychometric network analysis. In Study 2, we
validated the final 33-item version of the Mint against the established
interoception questionnaires, demonstrating its robust structure and
predictive validity for a range of clinically-relevant outcomes.
